---
title: "Compare Inventors Between Search Sets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compare Inventors Between Search Sets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
bibliography: citations.bib
csl: apa.csl
link-citations: true
---
  
*Built with R 
`r getRversion()`*
  
***

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5,
  fig.width = 8.84,
  dev = "CairoSVG",
  fig.ext = "svg"
)
library(uspto)
library(knitr)
outDir <- "eda/searches/"
if (!dir.exists(outDir)) outDir <- "../eda/searches/"
```

```{css, echo = FALSE}
td {white-space: nowrap}
```

This example explores the inventor-focused search set created by [example_inventor_sexing.R](https://github.com/uva-bi-sdad/uspto/blob/main/eda/example_inventor_sexing.R).

The files in produces are in the [eda/searches](https://github.com/uva-bi-sdad/uspto/tree/main/eda/searches) directory:
```r
outDir <- "eda/searches/"
```

# Load Data

First, we can load in the original search results:
```{r}
search_pharmaceutical <- read.csv(xzfile(paste0(outDir, "pharmaceutical.csv.xz")))
search_mechanical <- read.csv(xzfile(paste0(outDir, "mechanical.csv.xz")))
```

The pharmaceutical search returned `r prettyNum(nrow(search_pharmaceutical), big.mark = ",")` results, and the mechanical search
returned `r prettyNum(nrow(search_mechanical), big.mark = ",")` results.

The files are split into separate search sets, which we'll combine:
```{r}
inventors <- rbind(
  # based on the search pharmaceutical.ab.
  read.csv(xzfile(paste0(outDir, "inventors_pharmaceutical.csv.xz"))),
  # based on the search F01?.cpcl.
  read.csv(xzfile(paste0(outDir, "inventors_mechanical.csv.xz")))
)

# ensure our working set is from the original searches
inventors <- inventors[inventors$applicationNumber %in% c(
  search_mechanical$applicationNumber, search_pharmaceutical$applicationNumber
), ]
```

This set includes information from `r prettyNum(length(unique(inventors$guid)), big.mark = ",")` patent applications, with
`r prettyNum(length(unique(inventors$fullname)), big.mark = ",")` inventors (going by unique full name anyway). Of those
applications, `r prettyNum(length(unique(inventors[inventors$search_set == "pharmaceutical", "guid"])), big.mark = ",")`
are from the pharmaceutical search, and
`r prettyNum(length(unique(inventors[inventors$search_set == "mechanical", "guid"])), big.mark = ",")` are from mechanical set.

# Set Assessment

This set of applications is meant to represent technology fields that differ in their proportion of
female inventors, where mechanical (such as machine tools and mechanical elements) has a particularly
low proportion, and pharmaceutical (along with biotechnology and areas of chemistry) has a higher
proportion [@pto19; @ipo19]. The field-specific subsets are only defined by their simple search terms,
however, so we'll want to see how representative of those fields these applications might actually be.

One way to get a feel for how representative our sets are might be to look at the distribution of
classifications within each:
```{r}
# make a subset that contains only one line per applications
applications <- inventors[!duplicated(inventors$guid), ]

# look at the overall UPC class of each application between sets
overall_classes <- substring(applications$classification, 4, 6)
classes <- table(overall_classes, applications$search_set)

class_highlights <- rbind(
  # most differing classes
  classes[order(classes[, 2] - classes[, 1])[c(1:10, (1:10) + nrow(classes) - 10)], ],
  # most overlapping classes
  classes[order(abs(classes[, 2] - classes[, 1]) - rowSums(classes))[1:10], ]
)
class_highlights <- data.frame(
  class = rownames(class_highlights),
  criteria = rep(c("differing", "overlapping"), c(20, 10)),
  class_highlights
)

## add descriptions
library(uspto)
class_info <- get_class_info(class_highlights$class, paste0(dirname(outDir), "/original/class_info"))
class_highlights$description <- vapply(class_info, "[[", "", "description")
kable(
  class_highlights,
  row.names = FALSE,
  col.names = c("Class", "Criteria", "Mechanical", "Pharmaceutical", "Description")
)
```

Some other application-level features might also affect the comparisons we might want to make between sets,
so we can see how balanced they are incidentally:
```{r}
applications$n_inventors <- tapply(inventors$guid, inventors$guid, length)[applications$guid]
summaries <- vapply(split(applications, applications$search_set), function(d) {
  c(
    "Proportion US" = mean(d$inventorCountry == "US"),
    "Proportion California" = mean(d[d$inventorCountry == "US", "inventorState"] == "CA"),
    "Mean Year" = mean(as.numeric(substring(d$date, 1, 4))),
    "Proportion Utility" = mean(d$category == "Utility"),
    "Proportion Small Business" = (1 - mean(d$business == "UNDISCOUNTED")),
    "Proportion Inventor First" = mean(d$first_inventor == "true"),
    "Mean Inventors Per Team" = mean(d$n_inventors),
    "Mean Time To classification" = mean(d$time_initial_classificaiton, na.rm = TRUE),
    "Mean Time To First Action" = mean(d$time_first_action, na.rm = TRUE),
    "Mean Examination Rounds" = mean(d[d$examination_rounds != 0, "examination_rounds"]),
    "Proportion Accepted" = mean(d$any_accepts),
    "Proportion Patented" = mean(d$status == "Patented Case", na.rm = TRUE)
  )
}, numeric(12))
summaries <- cbind(summaries, summaries[, 1] - summaries[, 2])
kable(summaries, digits = 3, col.names = c("Mechanical", "Pharmaceutical", "Mechanical - Pharmaceutical"))
```

# Inventor Sexing

The only inventor information included in USPTO data is name, country, states (in some countries),
and city. We are particularly interested in looking at differences in the sex-distribution
of inventors between technology areas, so we used that inventor information to assign sex.
For some comparison, we used 3 basic methods with several different source:

- Historical or public name-sex datasets:
  - `sex_in_country_wgnd` and `prop_fem_wgnd` [with some additional sources\; @wgnd]
  - `prob_fem_skydeck` [@skydeck]
  - `prob_fem_usssa` [@ssa]
  - `sex_usuk` [@R-GenderInfer]
  - `prob_fem_ssa`, `prob_fem_ipums`, and `prob_fem_napp` [@R-gender]
- Social media profiles:
  - `prob_fem_fb` [from leaked account details\; @fb]
  - `prob_fb_scraped` [scraped from profiles\; @fbs]
- Search cues:
  - `prob_fem_search` ([guess_sex.R](https://github.com/miserman/lusilab/blob/main/R/guess_sex.R))

First, we can just look at the correlation between sources:
```{r}
prob_cols <- grep("prob_fem", colnames(inventors), fixed = TRUE, value = TRUE)
cors <- cor(inventors[, prob_cols], use = "pairwise.complete.obs")
rownames(cors) <- paste0("(", seq_along(prob_cols), ") ", prob_cols)
colnames(cors) <- paste0("(", seq_along(prob_cols), ")")
kable(cors, digits = 3)
```

The World Gender Name Dictionary has the broadest coverage and widest range of source,
so we could also treat that as the best guess, and look at accuracy of the others:
```{r}
# convert probabilities to predictions
sex_predictions <- inventors[, prob_cols]
sex_predictions[is.na(sex_predictions)] <- .5
sex_predictions[sex_predictions == .5] <- "U"
sex_predictions[inventors[, prob_cols] > .5] <- "F"
sex_predictions[inventors[, prob_cols] < .5] <- "M"
inventors[, sub("prob_fem", "sex", prob_cols, fixed = TRUE)] <- sex_predictions
sex_cols <- grep("sex_", colnames(inventors), fixed = TRUE, value = TRUE)

# then get accuracy to WGND in country
prediction_summaries <- data.frame(
  Accuracy = colMeans(vapply(
    inventors[, sex_cols], "==", logical(nrow(inventors)), inventors$sex_in_country_wgnd
  )),
  Percent_Determinate = colMeans(vapply(inventors[, sex_cols], "!=", logical(nrow(inventors)), "U"))
)
kable(prediction_summaries[order(-prediction_summaries$Accuracy),], digits = 3)
```

# Sex Differences

With the potential limitations of our sample and inventor sexing methods in mind, 
we can look at differences between sex-based inventor groups:
```{r}
inventors$n_inventors <- tapply(inventors$guid, inventors$guid, length)[inventors$guid]
summaries_sex <- vapply(split(inventors, inventors$sex_in_country_wgnd), function(d) {
  c(
    "Pharmaceutical Set" = mean(d$search_set == "pharmaceutical"),
    "Proportion US" = mean(d$inventorCountry == "US", na.rm = TRUE),
    "Proportion California" = mean(d[
      !is.na(d$inventorCountry) & d$inventorCountry == "US", "inventorState"
    ] == "CA"),
    "Mean Year" = mean(as.numeric(substring(d$date, 1, 4))),
    "Proportion Utility" = mean(d$category == "Utility"),
    "Proportion Small Business" = (1 - mean(d$business == "UNDISCOUNTED")),
    "Proportion Inventor First" = mean(d$first_inventor == "true"),
    "Mean Inventors Per Team" = mean(d$n_inventors),
    "Mean Time To classification" = mean(d$time_initial_classificaiton, na.rm = TRUE),
    "Mean Time To First Action" = mean(d$time_first_action, na.rm = TRUE),
    "Mean Examination Rounds" = mean(d[d$examination_rounds != 0, "examination_rounds"]),
    "Proportion Accepted" = mean(d$any_accepts),
    "Proportion Patented" = mean(d$status == "Patented Case", na.rm = TRUE)
  )
}, numeric(13))
kable(data.frame(
  Female = as.numeric(summaries_sex[, 1, drop = FALSE]),
  Male = summaries_sex[, 2],
  "Female - Male" = summaries_sex[, 1] - summaries_sex[, 2],
  Unknown = summaries_sex[, 3],
  check.names = FALSE
), digits = 3)
```

# References
