---
title: "Analyze inventor features using PatentsView data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyze inventor features using PatentsView data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

*Built with R 
`r getRversion()`*

***

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5,
  fig.width = 8.84,
  dev = "CairoSVG",
  fig.ext = "svg"
)
library(uspto)
library(knitr)
outDir <- "../patentsview/"
if (!dir.exists(outDir)) outDir <- "../../patentsview/"
```

```{css, echo = FALSE}
td {white-space: nowrap}
```

This example looks at a set of patent applications based on an initial seed set.

## Setup

Before starting, we'll need to load the package, and point to a directory where we'd like things saved:
```r
# install if needed: remotes::install_guthub("uva-bi-sdad/uspto")
library(uspto)
outDir <- "../patentsview/"
```

## Which patent categories have most and fewest female inventors?

The first step toward answering this question is to join the table listing inventor
details (which includes a name-based sex prediction) with that associating inventors
with patents:
```{r}
# start with inventors
inventors <- as.data.frame(download_patentsview_bulk("inventor", outDir))

## focus on just sexed inventors
inventors <- inventors[inventors$attribution_status == 1, ]
rownames(inventors) <- inventors$id

# add patent IDs
patent_inventor <- download_patentsview_bulk("patent_inventor", outDir)
patent_inventor <- patent_inventor[patent_inventor$inventor_id %in% inventors$id, ]
inventors <- cbind(inventors[patent_inventor$inventor_id, -1], patent_inventor)

# count inventors predicted to be female in each patent
female_inventors <- tapply(inventors$male_flag == 0, inventors$patent_id, sum)
```

Now we can add patent category information, to get breakdowns of classes by inventor sex.
There are multiple classification schemes, but we can start with the
World Intellectual Property Organization (WIPO) technology fields for a high-level overview:
```{r}
library(Matrix)

# this makes a patent x WIPO category matrix
categories_wipo <- patentsview_class_matrix("wipo", paste0(outDir, "wipo_matrix.rds"), dir = outDir)
dim(categories_wipo)

# which we can join to the inventors matrix
inventors_wipo <- female_inventors[names(female_inventors) %in% rownames(categories_wipo)]
inventors_wipo <- cbind(as.numeric(inventors_wipo), categories_wipo[names(inventors_wipo), ])

# and get category breakdowns by inventor sex
wipo_summary <- data.frame(
  Any_Female = colSums(inventors_wipo[inventors_wipo[, 1] != 0, -1] != 0),
  No_Female = colSums(inventors_wipo[inventors_wipo[, 1] == 0, -1] != 0)
)
wipo_summary$Any_Female_Proportion <- wipo_summary$Any_Female / rowSums(wipo_summary)
wipo_summary <- wipo_summary[order(-wipo_summary$Any_Female_Proportion), ]

# add category titles
wipo_field <- as.data.frame(download_patentsview_bulk("wipo_field", outDir))
rownames(wipo_field) <- wipo_field$id
rownames(wipo_summary) <- wipo_field[rownames(wipo_summary), "field_title"]

kable(wipo_summary, col.names = gsub("_", " ", colnames(wipo_summary), fixed = TRUE))
```

For a more refined classification, we might look at the United States Patent Classification (USPC),
which is most closely related to examination process:
```{r}
# get the patent x USPC category matrix
categories_uspc <- patentsview_class_matrix(
  "uspc_current", paste0(outDir, "uspc_current_matrix.rds"),
  dir = outDir
)
dim(categories_uspc)

# join to the inventors matrix
inventors_uspc <- female_inventors[names(female_inventors) %in% rownames(categories_uspc)]
inventors_uspc <- cbind(as.numeric(inventors_uspc), categories_uspc[names(inventors_uspc), ])

# and get category breakdowns by inventor sex
uspc_summary <- data.frame(
  Any_Female = colSums(inventors_uspc[inventors_uspc[, 1] != 0, -1] != 0),
  No_Female = colSums(inventors_uspc[inventors_uspc[, 1] == 0, -1] != 0)
)
uspc_summary$Any_Female_Proportion <- uspc_summary$Any_Female / rowSums(uspc_summary)
uspc_summary <- uspc_summary[order(-uspc_summary$Any_Female_Proportion), ]

# add category titles
uspc_field <- as.data.frame(download_patentsview_bulk("mainclass_current", outDir))
rownames(uspc_field) <- uspc_field$id
uspc_summary$Title <- uspc_field[rownames(uspc_summary), "title"]

kable(uspc_summary, col.names = gsub("_", " ", colnames(uspc_summary), fixed = TRUE))
```
