---
title: "WIPO Contrast"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WIPO Contrast}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

*Built with R 
`r getRversion()`*

***

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5,
  fig.width = 8.84,
  dev = "CairoSVG",
  fig.ext = "svg"
)
library(uspto)
library(knitr)
library(Matrix)
library(splot)
outDir <- "../patentsview/"
if (!dir.exists(outDir)) outDir <- "../../patentsview/"
```

```{css, echo = FALSE}
.table {margin-bottom: 40px}
td {white-space: nowrap}
```

From the [previous example](https://uva-bi-sdad.github.io/uspto/articles/example-patentsview_inventors.html)
we can see that inventors of different sex groups seem to be represented differently by classes of
patents. Here, we'll focus on two of the World Intellectual Property Organization (WIPO)
classes: Pharmaceuticals, which has a particularly high allocation of female-assigned inventors,
and Civil Engineering, which has a particularly high allocation of male-assigned inventors.

```r
library(uspto)
outDir <- "../patentsview/"
```

We'll start by collecting a unified set of patents that have either of the focal WIPO classes:
```{r}
contrast_inventors_file <- paste0(outDir, "contrast_inventors.csv.xz")
if (file.exists(contrast_inventors_file)) {
  contrast_inventors <- vroom::vroom(contrast_inventors_file, show_col_types = FALSE)
} else {
  # 1. identify patents that fall exclusively into one of the two focal WIPO classes
  wipo_fields <- download_patentsview_bulk("wipo_field", outDir)
  wipo_fields <- c(
    pharma = wipo_fields[wipo_fields$field_title == "Pharmaceuticals", "id", drop = TRUE],
    civil = wipo_fields[wipo_fields$field_title == "Civil engineering", "id", drop = TRUE]
  )
  wipo_matrix <- patentsview_class_matrix("wipo", paste0(outDir, "wipo_matrix.rds"), dir = outDir)
  wipo_pharma <- names(which(
    wipo_matrix[, wipo_fields[["pharma"]]] != 0 & wipo_matrix[, wipo_fields[["civil"]]] == 0
  ))
  wipo_civil <- names(which(
    wipo_matrix[, wipo_fields[["civil"]]] != 0 & wipo_matrix[, wipo_fields[["pharma"]]] == 0
  ))
  wipo_targets <- c(wipo_pharma, wipo_civil)

  # 2. identify the inventors associated with those patents
  patent_inventors <- download_patentsview_bulk(
    "patent_inventor", outDir,
    make_db = TRUE,
    partition = list(series_code = function(d) substr(d$patent_id, 1, 2))
  )
  contrast_inventors <- dplyr::collect(dplyr::filter(
    patent_inventors, patent_id %in% wipo_targets
  ))
  contrast_inventors$n_inventors <- as.numeric(tapply(
    contrast_inventors$inventor_id, contrast_inventors$patent_id, length
  )[contrast_inventors$patent_id])

  # 3. combine patent and inventor data
  patents <- download_patentsview_bulk("patent", outDir, make_db = TRUE)
  patents_target <- as.data.frame(dplyr::collect(dplyr::select(dplyr::filter(
    patents, as.character(number) %in% wipo_targets
  ), number, type, date, num_claims, withdrawn, kind)))
  rownames(patents_target) <- patents_target$number
  contrast_inventors <- cbind(contrast_inventors, patents_target[contrast_inventors$patent_id, -1])
  contrast_inventors$WIPO_class <- c(
    "FALSE" = "Civil Engineering", "TRUE" = "Pharmaceuticals"
  )[as.character(contrast_inventors$patent_id %in% wipo_pharma)]

  # 4. add assignee data
  patent_assignees <- download_patentsview_bulk(
    "patent_assignee", outDir,
    make_db = TRUE,
    partition = list(series_code = function(d) substr(d$patent_id, 1, 2))
  )
  assignees <- dplyr::collect(dplyr::filter(
    patent_assignees, patent_id %in% contrast_inventors$patent_id
  ))
  assignees <- download_patentsview_bulk("rawassignee", outDir, make_db = TRUE)
  assignees_target <- dplyr::collect(dplyr::filter(
    assignees,
    as.character(patent_id) %in% contrast_inventors$patent_id & sequence == 0 & !is.na(organization)
  ))
  contrast_inventors$assignee_id <- structure(
    assignees_target$assignee_id,
    names = assignees_target$patent_id
  )[contrast_inventors$patent_id]
  contrast_inventors$assignee_type <- structure(
    assignees_target$type,
    names = assignees_target$patent_id
  )[contrast_inventors$patent_id]
  contrast_inventors$assignee_organization <- structure(
    assignees_target$organization,
    names = assignees_target$patent_id
  )[contrast_inventors$patent_id]

  # 5. add location information
  locations <- download_patentsview_bulk("location", outDir, make_db = TRUE)
  locations_target <- as.data.frame(dplyr::collect(dplyr::filter(
    locations, id %in% contrast_inventors$location_id
  )))
  locations_target[is.na(locations_target$country), "country"] <- "US"
  locations_target <- locations_target[, !colnames(locations_target) %in% c("state", "county")]
  rownames(locations_target) <- locations_target$id
  contrast_inventors <- contrast_inventors[contrast_inventors$location_id %in% locations_target$id, -4]
  contrast_inventors <- cbind(contrast_inventors, locations_target[contrast_inventors$location_id, -1])
  contrast_inventors$n_countries <- as.numeric(tapply(
    contrast_inventors$country, contrast_inventors$patent_id, function(cs) length(unique(cs))
  )[contrast_inventors$patent_id])

  # 6. add sex predictions
  inventor_sex <- vroom::vroom(
    paste0(outDir, "inventor_sex.csv.xz"),
    show_col_types = FALSE
  )
  contrast_inventors$prob_fem <- structure(
    inventor_sex$prob_fem,
    names = inventor_sex$id
  )[contrast_inventors$inventor_id]
  contrast_inventors$pred_fem_patentsview <- structure(
    inventor_sex$pred_fem_patentsview,
    names = inventor_sex$id
  )[contrast_inventors$inventor_id]

  vroom::vroom_write(contrast_inventors, contrast_inventors_file, ",")
}
```

And get a high-level feel for the set:
```{r}
# sex breakdowns
## PatentsView assignments
contrast_inventors$sex_patentsview <- "Unknown"
contrast_inventors$sex_patentsview[contrast_inventors$pred_fem_patentsview == 1] <- "Female"
contrast_inventors$sex_patentsview[contrast_inventors$pred_fem_patentsview == 0] <- "Male"
sex_patentsview <- table(contrast_inventors$sex_patentsview, contrast_inventors$WIPO_class)
sex_patentsview <- sweep(sex_patentsview, 2, colSums(sex_patentsview), "/") * 100
kable(sex_patentsview, digits = 3, caption = "Sex Breakdown - PatentsView")
## Aggregate assignments
contrast_inventors$sex_aggregate <- "Unknown"
contrast_inventors$sex_aggregate[contrast_inventors$prob_fem > .5] <- "Female"
contrast_inventors$sex_aggregate[contrast_inventors$prob_fem < .5] <- "Male"
sex_aggregate <- table(contrast_inventors$sex_aggregate, contrast_inventors$WIPO_class)
sex_aggregate <- sweep(sex_aggregate, 2, colSums(sex_aggregate), "/") * 100
kable(sex_aggregate, digits = 3, caption = "Sex Breakdown - Aggregate")
## Combined -- defaults to PatentsView, falls back on aggregate
contrast_inventors$pred_female <- contrast_inventors$pred_fem_patentsview == 1 |
  (contrast_inventors$pred_fem_patentsview != 0 & contrast_inventors$prob_fem > .5)
contrast_inventors$pred_male <- contrast_inventors$pred_fem_patentsview == 0 |
  (contrast_inventors$pred_fem_patentsview != 1 & contrast_inventors$prob_fem < .5)
contrast_inventors$pred_sex <- "Unknown"
contrast_inventors$pred_sex[contrast_inventors$pred_female] <- "Female"
contrast_inventors$pred_sex[contrast_inventors$pred_male] <- "Male"
sex_combined <- table(contrast_inventors$pred_sex, contrast_inventors$WIPO_class)
sex_combined <- sweep(sex_combined, 2, colSums(sex_combined), "/") * 100
kable(sex_combined, digits = 3, caption = "Sex Breakdown - Combined")
# patent-level summary
contrast_inventors$patent_id <- as.character(contrast_inventors$patent_id)
unique_patents <- contrast_inventors[!duplicated(contrast_inventors$patent_id), ]
unique_patents$any_female <- tapply(
  contrast_inventors$pred_female, contrast_inventors$patent_id, any
)[unique_patents$patent_id]
nrow(unique_patents)

kable(rbind(
  "Countries Per Patent" = summary(unique_patents$n_countries),
  "Inventors Per Patent" = summary(unique_patents$n_inventors)
), digits = 3)
```

# Class Breakdown

WIPO classes are defined by sets of International Patent Classification classes,
which align with Cooperative Patent Classification (CPC) classes,
so we can take a finer look at our contrasts by breaking out those classes:

```{r}
cpc_matrix_file <- paste0(outDir, "contrast_cpc_matrix.rds")
if (file.exists(cpc_matrix_file)) {
  cpc_matrix <- readRDS(cpc_matrix_file)
} else {
  cpc <- download_patentsview_bulk("cpc_current", outDir, make_db = TRUE)
  cpc_matrix <- patentsview_class_matrix(dplyr::compute(dplyr::filter(
    cpc, patent_id %in% contrast_inventors$patent_id
  )), cpc_matrix_file, dir = outDir)
}

cpc_group <- as.data.frame(download_patentsview_bulk("cpc_group", outDir))
rownames(cpc_group) <- cpc_group$id

# see which CPC classes best represent each focal WIPO class
wipo_cpc <- vapply(c("Civil Engineering", "Pharmaceuticals"), function(class) {
  ids <- unique_patents$patent_id[unique_patents$WIPO_class == class]
  ids <- ids[ids %in% rownames(cpc_matrix)]
  colSums(cpc_matrix[ids, ] != 0)
}, cpc_matrix[1, ])
wipo_cpc <- as.data.frame(sweep(wipo_cpc, 2, table(unique_patents$WIPO_class), "/") * 100)
wipo_cpc$Difference <- wipo_cpc$Pharmaceuticals - wipo_cpc$`Civil Engineering`
wipo_cpc$Title <- cpc_group[rownames(wipo_cpc), "title"]
wipo_cpc <- wipo_cpc[order(-wipo_cpc$Difference), ]
kable(
  wipo_cpc[c(1:20, seq_len(20) + nrow(wipo_cpc) - 20), ],
  digits = 3, caption = "Percent of focal patents in which each CPC class appears"
)
```

We can also look at inventor sex breakdowns within CPC classes to see which CPC classes might be driving
the WIPO class inventor sex differences:
```{r}
sex_cpc <- vapply(c("Female", "Male"), function(sex) {
  ids <- contrast_inventors$patent_id[contrast_inventors$pred_sex == sex]
  ids <- ids[ids %in% rownames(cpc_matrix)]
  colSums(cpc_matrix[ids, ] != 0)
}, cpc_matrix[1, ])
sex_cpc <- as.data.frame(sweep(
  sex_cpc, 2, table(contrast_inventors$pred_sex, exclude = "Unknown"), "/"
) * 100)
sex_cpc$Difference <- sex_cpc$Female - sex_cpc$Male
sex_cpc$`Class Difference` <- wipo_cpc[rownames(sex_cpc), "Difference"]
sex_cpc$Title <- cpc_group[rownames(sex_cpc), "title"]
sex_cpc <- sex_cpc[order(-sex_cpc$Difference), ]
kable(
  sex_cpc[c(1:20, seq_len(20) + nrow(sex_cpc) - 20), ],
  digits = 3, caption = "Percent of inventors in each sex group on a patent with the given class"
)
```

# Inventor Sex Distribution Over Time

Frequencies of patent categories change over time:

```{r}
unique_patents$year <- substr(unique_patents$date, 1, 4)
wipo_trends <- as.data.frame(table(unique_patents$WIPO_class, unique_patents$year))
colnames(wipo_trends) <- c("WIPO_Class", "Year", "Patents")
library(splot)
splot(Patents ~ Year * WIPO_Class, wipo_trends, line.type = "b", title = FALSE)
```

Frequencies of new inventors by assigned sex also changes over time:

```{r}
contrast_inventors$year <- substr(contrast_inventors$date, 1, 4)
contrast_inventors <- contrast_inventors[order(contrast_inventors$date), ]
unique_inventors <- contrast_inventors[!duplicated(contrast_inventors$inventor_id), ]
sex_trends <- as.data.frame(table(unique_inventors$pred_sex, unique_inventors$year))
colnames(sex_trends) <- c("Assigned_Sex", "Year", "Inventors")
splot(Inventors ~ Year * Assigned_Sex, sex_trends, line.type = "b", title = FALSE)
```

How much might a difference in sex distribution between classes be due to time? That is,
we might expect to see more new inventor participation in the dominant or rising class
regardless of sex, which could shift to sex distributions between classes.

To get at this, we might look at the distribution of first-seen inventors by sex group between classes over time:

```{r}
wipo_sex_yearly <- as.data.frame(t(vapply(split(
  unique_inventors[, c("WIPO_class", "pred_sex")], unique_inventors$year
), function(d) {
  s <- table(d$pred_sex, d$WIPO_class, exclude = "Unknown")
  s <- s / rowSums(s) * 100
  s[, 2] - s[, 1]
}, numeric(2))))
wipo_sex_yearly$Year <- rownames(wipo_sex_yearly)
splot(
  wipo_sex_yearly[, 1:2] ~ wipo_sex_yearly$Year,
  line.type = "b",
  title = "Sex Distribution Ballance Between WIPO Classes", labx = "Year",
  laby = "Percent Difference (Pharmaceuticals - Civil Engineering)"
)
```

Here we can see that first-time female-assigned inventors have consistently been more shifted toward
pharmaceuticals than first-time male-assigned inventors, and that in both groups there has been
a fairly consistent trend toward pharmaceuticals. Though it is interesting to see that in the early
years, first-time female-assigned inventors were more evenly split between classes, and
later on, first-time male-assigned inventors were more evenly split between classes, though there
is a more recent trend in the original direction.
